---
title: "FARS Functions in the week4 Package"
author: "Steven Shum"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FAR Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
library(tidyr)
```

## Introduction

The US National Highway Traffic Safety Administration's Fatality Analysis Reporting System (FARS) is a nationwide census providing American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. The week4 package provides functions which can be used to load the data into R for basic analysis and mapping. These are:
(1) fars_read
(2) make_filename
(3) fars_read_years
(4) fars_summarise_years
(5) fars_map_state

## fars_read and make_filename

The fars_read function reads the FARS data file and loads it into a data frame in R for subsequent analysis. If the file does not exist, it returns a file does not exist message. 

The FARS data files share a common format - accident_<year>.csv.bz2. The make_filename function provides a easy way to return the full filename to be entered by simply keying in the year. 

```{r}
dat <- fars_read(make_filename(2013))
```

## fars_read_years

The fars_read_years function loads files from multiple years and returns the months and years data from these files. Input argument is year or a list of years (e.g. 2013).

```{r}
fars_read_years(c(2013, 2014))
```

## fars_summarize_years

The fars_summarize_years function summarises the number of accidents by year and month. Input argument is year or a list of years (e.g. 2013).

```{r}
fars_summarize_years(c(2013, 2014))
```

## fars_map_state

The fars_map_state command provides a visualisation of the location of accidents in a given state in a given year. State should be entered as a 1 or 2-digit code (1 to 56). Below shows a sample plot for state = 20 in year 2013. 

```{r}
fars_map_state (20, 2013)
```

